---
title: "Practical Machine Learning Project"
author: "Gabriel Sotomayor"
date: "08-06-2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r enviroment}
set.seed(1920)
setwd("C:/Users/gabri/OneDrive/Coursera/Practical Machine Learning")
library(caret)
library(randomForest)

```
# Introduction and summary

On this work, we are using accelerometer data from 6 subjects to predict if they are doing barbell lifts on correct or incorrect ways.
To do that we cleaned the data, divided it on training and test data set and fit a random forest model. The model had an out of sample accuracy of 99.81% per cent on the testing data set and predicted correctly 100% per cent of the 20 extra cases provided to test the model.

# Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self-movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behaviour, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Data

We loaded the training and testing datasets provided by the course, divided the training data set on training and a test data set. Then we cleaned the data by removing the variables which had zero or near-zero variance, the variables with mostly NA cases, and the identification variables.

```{r data}
training <- read.csv(url("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"))
testing  <- read.csv(url("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"))

# create a partition with the training dataset 
inTrain  <- createDataPartition(training$classe, p=0.7, list=FALSE)
Traindata <- training[inTrain, ]
Testdata  <- training[-inTrain, ]
dim(Testdata)

zerv <- nearZeroVar(Traindata)
Traindata <- Traindata[, -zerv]
Testdata  <- Testdata[, -zerv]

RMNA    <- sapply(Traindata, function(x) mean(is.na(x))) > 0.95
Traindata <- Traindata[, RMNA ==FALSE]
Testdata <- Testdata[, RMNA ==FALSE]

Traindata<- Traindata[, -(1:5)]
Testdata  <- Testdata[, -(1:5)]
dim(Testdata)
```

# Model building and testing
We applied a random forest model to model the regressions
```{r random forest model}
set.seed(1920)
RF <- trainControl(method="cv", number=3, verboseIter=FALSE)
rforest <- train(classe ~ ., data=Traindata, method="rf",
                          trControl=RF)
rforest$finalModel
```

Then we used the training subset to create a prediction and then measure its accuracy. It predicted correctly on over 99% of the cases, so we decided to use this model to the final exercise of prediction.
```{r random forest model test}
# prediction on Test dataset
predrforest <- predict(rforest, newdata=Testdata)
cmrf <- confusionMatrix(predrforest, Testdata$classe)
cmrf
```

# Applying prediction model

To finish the exercise, we applied the model to predict on the testing data, with 100% per cent of accuracy.
```{r testing prediction}
predtesting <- predict(rforest, newdata=testing)
predtesting
```